%============================================================
% HALAMAN ABSTRACT
%============================================================
\cleardoublepage
\chapter*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT}
\thispagestyle{fancy}

\textit{The National Health Insurance (JKN) program applies the Indonesian Case-Based Groups (INA-CBG) payment mechanism for advanced healthcare facilities, making claim values highly dependent on diagnoses, procedures, severity levels, and grouping outcomes. This condition creates a risk of fraud or upcoding, which may lead to financial loss and complicate claim verification. The main challenge in this domain is how to identify high-risk claims in a fast, measurable, and explainable manner, especially when actual fraud labels are unavailable at the claim level.}

\textit{This study proposes a machine-learning-based approach for detecting fraud/upcoding risk indications using a claim-centric strategy, where each claim is treated as the main unit of analysis and enriched with administrative, clinical, cost, secondary diagnosis, and service history information. To address the absence of actual fraud labels, the study constructs initial targets through pseudo-labeling with Isolation Forest, followed by supervised multi-model benchmarking. The proposed pipeline also includes feature engineering, branch-based feature selection, threshold tuning, and both global and local interpretability.}

\textit{Experiments were conducted on 1,699,219 integrated FKRTL claims. The best configuration was achieved by XGBoost with the VIF branch and a threshold of 0.80, yielding a ROC-AUC of 0.999876, PR-AUC of 0.996046, F1-score of 0.952511, precision of 0.913917, recall of 0.994507, and MCC of 0.951883 against the pseudo-label target. Interpretability analysis indicates that cost-related features, grouping tariff, non-capitation service history, special CMG, and tariff ratios are key drivers of the risk score. Therefore, the proposed system has strong potential to support claim verification by prioritizing high-risk claims rather than issuing final fraud judgments.}

\vspace{1cm}
\noindent\textbf{Keywords:} Fraud Detection, Upcoding, INA-CBG, Machine Learning, Claim-Centric
